\documentclass[11pt,a4paper]{report}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{mathrsfs}
\usepackage{textcomp}

\newcommand{\zm}{ZM}
\newcommand{\zsystem}{Zebra-machine (\zm)}
\newcommand{\var}[1]{`#1}
\newcommand{\zconst}[1]{\textbf{#1}}
\newcommand{\zvar}[1]{\textit{\var{#1}}}
\newcommand{\ztuple}[1]{(#1)}
\newcommand{\zneg}[1]{\textlnot \ztuple{#1}}
\newcommand{\zquery}[1]{?\ztuple{#1}}
\newcommand{\zunify}[3]{#1 \texttimes #2 = #3}
\newcommand{\zexample}[1]{\begin{center} #1 \end{center}}

\newcommand{\unify}{\otimes}

\begin{document}

\title{The Great Book of Zebra}
\author{The Zebra Project}
\date{\today}
\maketitle

\chapter*{Preface}

This book is a collaborative work from the \href{zebrajs}{https://github.com/fsvieira/zebrajs} project community
and everyone is invited to participate.

The list of contributors is at the contributors section~\ref{sec:contributors} and your name can be there too :D.

This is a work in progress.

\tableofcontents

\newpage
\section{Introduction}

This is the official book of \zsystem. Here you will find anything you need to understand in deep the \zm ,
the book covers both theoretical and practical definitions.

\zsystem\ is a logical symbolic computation query system, given a set of computational definitions it will
answer questions about them, therefor \zm is better suited for software validation and constrain satisfaction problems.

\chapter{The \zsystem}

On this chapter we will present an informal and formal defintions of \zsystem, we will try to explain how it is defined and 
how it works.

\section{Informal Definition}

Zebrajs is a logical symbolic computation query system, given a set of computational definitions it will answer questions about them, 
therefor Zebrajs is better suited for software validation and constraint satisfaction problems.

Zebrajs system consists of two parts the definitions and the query, both parts share the same language of zebra-system terms, 
which is defined by a certain formal syntax, and a set of transformation rules.

The zebra language is very simple, it only has constants, variables, tuples and negation.

In this section we will try to explain how it works using simple examples.

\subsection{Definitions, Tuples and constants}
    \paragraph{Definitions} are considered facts they are always tuples and always global.

    \zexample{(\zconst{yellow} \zconst{blue})}
    
    In this example we define a tuple with only constants, constants don't need to be defined anyhere as we consider that all constants exists.
    Because this tuple is defined on top level it is considered a definition and its considered to be true, similiar to an axiom in math.

    \zexample{(\zconst{yellow} (\zconst{blue}))}
    
    In this example the outer most tuple is considered a definition, iff its inner tuples are unfiable with a definition, because of this 
    inner tuples are considered to be queries and not facts.
    
    Two tuples are unifiable if both contains the same number of elements and all its elements are unfiable.
    Two constants are unfiable if they are the same (eg. \zunify{yellow}{yellow}{yellow}).

\subsection{Variables}
    \paragraph{Variables} can only appear inside of tuples and they can unify with anything (tuples, constants and variables).
    
    A variable is represented as \zvar{v}, where v is a variable. We can also use anonimous variables like this \zvar{}.

    For example:
    \zexample{(\zvar{name} \zvar{age})}

    This tuple contains two variables, the \zvar{name} and \zvar{age}.

    The variable scope is associated with the top tuple, for example:
    \zexample{
        \begin{enumerate}
            \item (\zvar{name} (\zvar{name} \zvar{age}))
            \item (\zvar{name})
        \end{enumerate}
    }

    On first tuple variable \zvar{name} is bound to the inner-tuples, but it is not 
    the same variable \zvar{name} that appears on the second tuple.

    We can also make anonimous variables, by not giving them any name, anonimous variables 
    are not bound to any other variables.
    \zexample{(\zvar{} \zvar{})}

\subsection{Queries}
    \paragraph{Queries} they can only appear at the top level, they are queries to definitions and so they can only be applied to tuples.
    A query as the form of \zquery{tuple}.
    
    For example:
        \zexample{
            \begin{enumerate}
                \item (\zconst{color} \zconst{blue})
                \item (\zconst{color} \zconst{yellow})
                \item (\zconst{color} \zconst{green})
                \item (\zconst{colors} \zconst{blue} \zconst{green})
                \item \zquery{\zconst{color} \zvar{}}
            \end{enumerate}
        }
    
    And the results of the query \zquery{\zconst{color} \zvar{}} are:
    \zexample{
        \begin{enumerate}
            \item (\zconst{color} \zconst{blue})
            \item (\zconst{color} \zconst{yellow})
            \item (\zconst{color} \zconst{green})
        \end{enumerate}
    }

    This means that query unifies with all 1, 2, 3 definitions, and fails to unify 
    with the 4 definition. 

\subsection{Negation}
    \paragraph{Negation} (\zneg{tuple}) can only appear inside of tuples and can only negate tuples.
    A tuple can contain multiple negations and they do not contribute for tuple length, they 
    are considered conditions for tuple success. Because of this negations are not considered for tuple unification 
    however they are unified with definitions so they can be evaluated to hold or not hold. 

    A negation works as a $\nexists$, for example:
    
    \zexample{
        \begin{enumerate}
            \item (\zconst{equal} \zvar{p} \zvar{p})
            \item (\zconst{not-equal} \zvar{p} \zvar{q} \zneg{\zconst{equal} \zvar{p} \zvar{q}})
        \end{enumerate}
    }
    
    In this example both tuples have length three, because negation are not considered for unification, they work as conditions.
    On the second tuple we define a not-equal negating the equal definition, this means that if \zvar{p} and \zvar{q} are equal
    than equal definition holds, and negation doesn't so not-equal also fails.
    
    If \zvar{p} and \zvar{q} are different than equal fails, negation will hold and not-equal will hold.
    
    \zvar{p} and \zvar{q} are only different if their values are different so not-equal will always fails if variables have no values, 
    this is because there is no garantee that \zvar{p} is not equal to \zvar{q}.
    
    This doesn't mean that our negations queries can only be performed with values, here is another example:
    
    \zexample{
        \begin{enumerate}
            \item \ztuple{\zconst{color} \zconst{yellow}}
            \item \ztuple{\zconst{color} \zconst{blue}}
            \item \ztuple{\zconst{color} \zconst{green}}
            \item \ztuple{\zconst{equal} \zvar{p} \zvar{p}}
            \item \ztuple{\zconst{not-equal} \zvar{p} \zvar{q} \zneg{\zconst{equal} \zvar{p} \zvar{q}}}
            \item \zquery{\zconst{not-equal} \ztuple{\zconst{color} \zvar{}} \ztuple{\zconst{color} \zvar{}}}
        \end{enumerate}
    }

    Ok, I cheated the query replaces variables with tuple values, but its only because we want to define a domain "colors". 
    So the result to the query should be pairs of distinct colors, like this:
    
    \zexample{
        \begin{enumerate}
            \item \ztuple{\zconst{not-equal} \ztuple{\zconst{color} \zconst{yellow}} \ztuple{\zconst{color} \zconst{blue}}}
            \item \ztuple{\zconst{not-equal} \ztuple{\zconst{color} \zconst{yellow}} \ztuple{\zconst{color} \zconst{green}}}
            \item \ztuple{\zconst{not-equal} \ztuple{\zconst{color} \zconst{blue}} \ztuple{\zconst{color} \zconst{yellow}}}
            \item \ztuple{\zconst{not-equal} \ztuple{\zconst{color} \zconst{blue}} \ztuple{\zconst{color} \zconst{green}}}
            \item \ztuple{\zconst{not-equal} \ztuple{\zconst{color} \zconst{green}} \ztuple{\zconst{color} \zconst{yellow}}}
            \item \ztuple{\zconst{not-equal} \ztuple{\zconst{color} \zconst{green}} \ztuple{\zconst{color} \zconst{blue}}}
        \end{enumerate}
    }

\section{Formal Definition}

As mentioned before \zm\ is a logical symbolic computation query system, and it consists of two parts
the definitions and the query, both parts share the same language of \zm\ terms, which is defined by
a certain formal syntax, and a set of transformation rules.

\subsection{\zm\ Language ($\mathbb{L}$)}
\paragraph{The \zm\ language ($\mathbb{L}$) of terms are defined as:}

\begin{enumerate}
\item $\mathbb{C} = \{c: \text{c is a constant} \}$,
\item $\mathbb{V} = \{v: \text{v is a variable} \}$,
\item $\mathbb{N}_{0} = \{\neg (p_{0} \ldots p_{n}): p_{k} \in \mathbb{C} \cup \mathbb{V}, k \geq 0 \wedge k \leq n \}$, where $\neg t$ is the negation of tuple $t$.
\item $\mathbb{L}_{0} = \{(p_{0} \ldots p_{n}) : p_{k} \in \mathbb{C} \cup \mathbb{V} \cup \mathbb{N}_{0}, k \geq 0 \wedge k \leq n \}$

\item $\mathbb{N}_{i} = \{\neg t: t \in \mathbb{L}_{i-1}\}$, $i > 0$
\item $\mathbb{L}_{i} = \{(p_{0} \ldots p_{n}) : p_{k} \in \mathbb{C} \cup \mathbb{V} \cup \mathbb{L}_{i-1} \cup \mathbb{N}_{i}, k \geq 0 \wedge k \leq n \}$, $i > 0$

\item $\mathbb{L} = \bigcup\limits_{i=0}^{\infty} \mathbb{L}_{i} \cup \{?t : t \in \mathbb{L}_{i} \}$, where $?t$ is a query and $t$ a tuple.
\end{enumerate}

% \begin{enumerate}
% \item $\mathbb{L}_{0}$ and $\mathbb{L}_{not}$ are sets of \zm\ terms.
% \item $c \in \mathbb{C}$: $\mathbb{C}$ is the set of terminal symbols called constants, $c$ is a terminal symbol. $c \in \mathbb{L}_{0}$.
% \item $\var{p} \in \mathbb{V}$: $\mathbb{V}$ is the set of variables, $p$ is a variable. $\var{p} \in \mathbb{L}_{0}$.
% \item $(p_{0} \ldots p_{n}) \in \mathbb{T}$: $\mathbb{T}$ is the set of tuples, $(p_{0} \ldots p_{n})$ its a n-tuple of \zm\ terms. $(p_{0} \ldots p_{n}) \in \mathbb{L}_{0}$.
% \item $\overline{p}$ can be any term that is not equal to $p$, $p \in \mathbb{L}_{0}$. $\overline{p} \in \mathbb{L}_{not}$.
% \item Nothing else is a \zm\ term.
% \item $\mathbb{L} = \mathbb{L}_{0} \cup \mathbb{L}_{not}$,
% \end{enumerate}


\subsection{\zm\ Operations}

\paragraph{The powerset of set $\mathbb{S}$ is expressed as $\mathrm{P}(\mathbb{S})$}.

\paragraph{Unification ($\unify$, binary operation)} defined as
\[
    \unify: \mathbb{L} \times \mathbb{L} \rightarrow \mathbb{L}
\]

and by the rules,

\[
    \unify: \mathbb{L}_{0} \times \mathbb{L}_{0} \rightarrow \mathbb{L}_{0}
\]

\begin{enumerate}
\item $Q \unify Q = Q$
    \subitem $Q$ unifies with itself, resulting on itself.

\item $\var{p} \unify Q = Q \iff \var{p} = Q$
\item $Q \unify \var{p} = Q \iff \var{p} = Q$
    \subitem note that if $Q \in \mathbb{V}$ then rule [2] will also apply, since both $p$ and $Q$ are the same the result is also the same.

\item $(p_{0} \ldots p_{n}) \unify (q_{0} \ldots q_{n}) \iff (p_{0} \unify q_{0} \ldots p_{n} \unify q_{n})$
    \subitem $(p_{0} \ldots p_{n})$ z-tuple only unifies with other z-tuple if they have same size and all sub \zm\ terms unify.

\end{enumerate}


\[
    \unify: \mathbb{L}_{0} \times \mathbb{L}_{not} \rightarrow \mathbb{L}_{0}
\]

\begin{enumerate}
\item $Q \unify \overline{p} = Q \iff p \neq Q$
\end{enumerate}


\[
    \unify: \mathbb{L}_{not} \times \mathbb{L}_{0} \rightarrow \mathbb{L}_{0}
\]

\begin{enumerate}
\item $\overline{p} \unify Q = Q \iff p \neq Q$
\end{enumerate}

\[
    \unify: \mathbb{L}_{not} \times \mathbb{L}_{not} \rightarrow \mathbb{L}_{0}
\]

\begin{enumerate}
\item $\overline{P} \unify \overline{Q} = \var{p} \iff \var{p} \unify \overline{P} \wedge \var{p} \unify \overline{Q}$
    \subitem where $\var{p}$ is a new variable.
    \subitem note that if $P = Q$, then $\var{p} \neq (P \unify Q)$.
\end{enumerate}

Anything else is not unifiable.

\paragraph{Substitution ($\mathcal{S}$, function)} defined as:

\begin{align}
    \mathcal{S}&: \mathbb{V} \times \mathbb{L} \times \mathbb{L} \rightarrow \mathbb{L}\\
    \mathcal{S}&(v, w, t) =
    \begin{cases}
        \hfill w \hfill &\text{, if } t = v,\\
        \hfill (\mathcal{S}(v, w, t_{0}) \ldots \mathcal{S}(v, w, t_{n})) \hfill & \text{, if } t = (t_{0} \ldots t_{n})\\
        \hfill t \hfill & \text{, otherwise}
    \end{cases}
\end{align}


\paragraph{The notation $\mathbb{T}_{p}$} is used to define the set containing all of $p$ inner tuples.

\subsection{\zm\ Computation}
\paragraph{A \zm\ computation is expressed as 4-tuple $(\sigma, \delta, q, \alpha)$ where:}

\begin{enumerate}
\item $\sigma$ is a set of terminal symbols (constants),
\item $\delta$ is a set of z-tuples (definitions),
\item $q$ is a z-tuple (query),
\item $\alpha$ is the set of possible computational answers to query $q$ based on $delta$ definitions.
\end{enumerate}

\paragraph{A definition} is a fact in the system. The inner tuples of a definition are considered and called queries, therefor
for a definition to be true all of its inner tuples/queries must also be true.

\paragraph{A query} is a question to the system that is true if and only if it unifies at least with
one definition.

\paragraph{Free and bound variables} on the context of a definition all definition variables are considered
to be bound to the definition, on the context of queries all variables are free.

Bound variables will be expressed as:
\[
    ([v_{0} \ldots v_{n}] p_{0} \ldots p_{k})
\]

where $v_{i}, 0 <= i <= n$ are the tuple bound variables and $p_{j}, 0 <= j <=k$ are the tuple terms.

ex:
\[
    ([a\ b\ c]\ (\var{a}\ \var{b})\ \var{c}\ \text{yellow})
\]

Bound information and notation is used on intermidiate computation steps, therefor they are not included on the $\zm$ language.


\paragraph{Variable renaming} its necessary to ensure that distinct variables with same name are not handled as being the same.

ex:
\[
    ([p]\ \var{p}\ \var{q}) \unify ([p\ q]\ \var{p}\ \var{q})
\]

The first tuple has only $\var{p}$ as bound variable and the second tuple as $\var{p}$ and $\var{q}$ declared as bound variables.
To ensure that this variables keep their meaning, we rename the bound variables like this:

\[
    ([p_{0}]\ \var{p_{0}}\ \var{q}) \unify ([p_{1}\ q_{1}]\ \var{p_{1}}\ \var{q_{1}})
\]

After unification we get

\[
    ([p_{0}\ p_{1}\ q_{1}]\ \var{p_{0}} \unify \var{p_{1}}\ \var{q} \unify \var{q_{1}})
\]


\paragraph{A computation} is done with the following steps:

\begin{enumerate}
    \item $\delta_{0} = \{ p_{0} \unify \ldots \unify p_{|x|} : \forall x \in \mathrm{P}(\delta) \wedge \forall p_{i} \in x, i=0 \ldots |x| \}$
    \item $\delta_{1} = \{ ([v_{0} \ldots v_{n}]\ p_{0}\ p_{k}): \forall (p_{0} \ldots p_{n}) \in \delta_{0})\}$
    \subitem where $v_{0} \ldots v_{n}$ are all (bound) variables found on $(p_{0} \ldots p_{n})$ tuple.
    \subitem $\delta_{1}$ is the set of definitions with bound variables.
    \item $\alpha = \{q \unify p :\forall p \in \delta_{1}\ \wedge \forall t \in \mathcal{T}_{q \unify q} \rightarrow (\sigma, \delta, t, \alpha) \}$
    \subitem the $\alpha$ set is the result of the unification of the query $q$ with all definitions $p$ and the resulting innter tuples are
    also a valid computation.
    \subitem a computational will stop if the set of inner tuples are empty.
    \subitem a computational may not stop for some cases.
\end{enumerate}

\begin{comment}
\section{Computing Examples}

\subsection{Defining a list}

\begin{enumerate}
\item $(\text{list})$
\item $(\text{list}\ \var{item}\ (\text{list}))$
\item $(\text{list}\ \var{item}\ (\text{list}\ \var{i}\ \var{tail}))$
\end{enumerate}

Ex:
\[
    (\sigma, \delta, q, \alpha ) = (\{\text{yellow}, \text{blue}\},
    \{
(\text{list}),
(\text{list}\ \var{item}\ (\text{list})),
(\text{list}\ \var{item}\ (\text{list}\ \var{i}\ \var{tail}))
    \}, q, \alpha)
\]

\begin{enumerate}
\item $q = (\text{list})$, query a empty list.
\subitem $(\text{list}) \unify (\text{list}) = (\text{list})$
\subitem all other definitions fail to unify because the number of elements in remaining definitions don't match query tuple.
\subitem $\alpha = {(\text{list})}$
\item $q = (\text{list}\ \text{yellow}\ (\text{list}\ \text{blue}\ (\text{list})))$
\subitem Starting with definition $\text{list}$,
\subitem $(\text{list}\ \text{yellow}\ (\text{list}\ \text{blue}\ (\text{list}))) \unify (\text{list})$, fail.
\subitem Next definition $(\text{list}\ \var{item}\ (\text{list}))$,
\subitem $(\text{list}\ \text{yellow}\ (\text{list}\ \text{blue}\ (\text{list}))) \unify (\text{list}\ \var{item}\ (\text{list}))[\text{bound}: \text{item} ]$
\subitem $(\text{list} \unify \text{list}\ \text{yellow} \unify \var{item}\ (\text{list}\ \text{blue}\ (\text{list})) \unify (\text{list}))[\text{bound}: \text{item}]$
\subitem fail to unify $(list)$ with $(\text{list}\ \text{blue}\ (\text{list}))$.
\subitem Next definition $(\text{list}\ \var{item}\ (\text{list}\ \var{i}\ \var{tail}))$,
\subitem $(\text{list}\ \text{yellow}\ (\text{list}\ \text{blue}\ (\text{list}))) \unify (\text{list}\ \var{item}\ (\text{list}\ \var{i}\ \var{tail}))[\text{bound}: \text{item}\ \text{i}\ \text{tail}]$
\subitem $(\text{list}\unify \text{list}\ \text{yellow} \unify \var{item}\ (\text{list}\ \text{blue}\ (\text{list})) \unify (\text{list}\ \var{i}\ \var{tail}))[\text{bound}: \text{item}\ \text{i}\ \text{tail}]$
\subitem $(\text{list}\ \text{yellow} (\text{list}\unify \text{list}\ \text{blue} \unify \var{i}\ (\text{list}) \unify \var{tail}))[\text{bound}: \text{item}\ \text{i}\ \text{tail}, \text{constrains}: \var{item} = \text{yellow}]$
\subitem $(\text{list}\ \text{yellow} (\text{list}\ \text{blue}\ (\text{list})))
[\text{bound}: \text{item}\ \text{i}\ \text{tail}, \text{constrains}: \var{item} = \text{yellow}\ \var{i}=\text{blue}\ \var{tail}=(\text{list})]$
\subitem the next step is to unify all sub-tuples that are not yet checked with definitions, this are: $(\text{list}\ \text{blue}\ (\text{list}))$ and $(\text{list}))$,
\subitem $(\text{list}\ \text{blue}\ (\text{list})) \unify (list)$, fail.
\subitem $(\text{list}\ \text{blue}\ (\text{list})) \unify (\text{list}\ \var{item}\ (\text{list}))[\text{bound}: item]$
\subitem $(\text{list}\unify \text{list}\ \text{blue}\unify \var{item}\ (\text{list}) \unify (\text{list}))[\text{bound}: item]$
\subitem $(\text{list} \text{blue} (\text{list})))[\text{bound}: item, \text{constrains}: \var{item}=\text{blue}]$, succeed.
\subitem $(\text{list}\ \text{blue}\ (\text{list})) \unify (\text{list}\ \var{item}\ (\text{list}\ \var{i}\ \var{tail}))$, fail.
\subitem Finally $(list)$ will only unify with $(list)$ definition, ending the process and resulting on $\alpha = \{ (\text{list}\ \text{yellow}\ (\text{list}\ \text{blue}\ (\text{list}))) \}$
\end{enumerate}

\subsection{What john likes?}

A simple example of querying facts.

We start with the definitions/facts:
\begin{enumerate}

\item $(\text{mary}\ \text{likes}\ \text{wine}\ \var{p})$
\subitem mary likes wine.

\item $(\text{mary}\ \text{likes}\ \text{john}\ \var{p})$
\subitem mary likes john.

\item $(\text{peter}\ \text{likes}\ \text{peter}\ \var{p})$
\subitem peter likes himself.

\item $(\text{john}\ \text{likes}\ \var{stuff}\ (\text{mary}\ \text{likes}\ \var{stuff}\ \var{p}))$
\subitem john likes everything that mary likes.

\item $(\text{john}\ \text{likes}\ \var{person}\ (\var{person}\ \text{likes}\ \text{wine}\ \var{p}))$
\subitem john likes anyone that likes wine.

\item \[
    (\text{john}\ \text{likes}\ \var{person}\ (\text{list}\ (\var{person}\ \text{likes}\ \var{person}\ \var{p}) (\text{list}\ (\text{notEqual}\ \var{person} \text{john})\ (\text{list}))))
\]
\subitem john likes anyone that likes themselves.

\end{enumerate}

\end{comment}

\section{Computing Examples (implementation)}

\subsection{Defining a list}

\begin{enumerate}
\item $(\text{list})$
\item $(\text{list}\ \var{item}\ (\text{list}))$
\item $(\text{list}\ \var{item}\ (\text{list}\ \var{i}\ \var{tail}))$
\end{enumerate}

Ex:
\[
    (\sigma, \delta, q, \alpha ) = (\{\text{yellow}, \text{blue}\},
    \{
(\text{list}),
(\text{list}\ \var{item}\ (\text{list})),
(\text{list}\ \var{item}\ (\text{list}\ \var{i}\ \var{tail}))
    \}, q, \alpha)
\]

\begin{enumerate}
\item $q = (\text{list})$, query a empty list.
\subitem $(\text{list}) \unify (\text{list}) = (\text{list} \unify \text{list}) = (\text{list})$
\subitem all other definitions fail to unify because the number of elements in remaining definitions don't match query tuple.
\subitem $\alpha = \{(\text{list})\}$
\item $q = (\text{list}\ \text{yellow}\ (\text{list}\ \text{blue}\ (\text{list})))$
\subitem Starting with definition $(\text{list})$,
\subitem $(\text{list}\ \text{yellow}\ (\text{list}\ \text{blue}\ (\text{list}))) \unify (\text{list})$, fail.
\subitem Next definition $(\text{list}\ \var{item}\ (\text{list}))$,
\subitem $(\text{list}\ \text{yellow}\ (\text{list}\ \text{blue}\ (\text{list}))) \unify ([item]\ \text{list}\ \var{item}\ (\text{list}))$
\subitem $([item]\ \text{list} \unify \text{list}\ \text{yellow} \unify \var{item}\ (\text{list}\ \text{blue}\ (\text{list})) \unify (\text{list}))$
\subitem fail to unify $(list)$ with $(\text{list}\ \text{blue}\ (\text{list}))$.
\subitem Next definition $(\text{list}\ \var{item}\ (\text{list}\ \var{i}\ \var{tail}))$,
\subitem $(\text{list}\ \text{yellow}\ (\text{list}\ \text{blue}\ (\text{list}))) \unify ([\text{item}\ \text{i}\ \text{tail}]\ \text{list}\ \var{item}\ (\text{list}\ \var{i}\ \var{tail}))$
\subitem $([\text{item}\ \text{i}\ \text{tail}]\ \text{list}\unify \text{list}\ \text{yellow} \unify \var{item}\ (\text{list}\ \text{blue}\ (\text{list})) \unify (\text{list}\ \var{i}\ \var{tail}))$
\subitem unify constants,
\subitem $([\text{item}\ \text{i}\ \text{tail}]\ \text{list}\ \text{yellow} \unify \var{item}\ (\text{list}\ \text{blue}\ (\text{list})) \unify (\text{list}\ \var{i}\ \var{tail}))$
\subitem $([\text{item}\ \text{i}\ \text{tail}]\ \text{list}\ \text{yellow} \unify \var{item}\ (\text{list} \unify \text{list}\ \text{blue}\ \unify \var{i}\ (\text{list}) \unify \var{tail}))$
\subitem unify constants,
\subitem $([\text{item}\ \text{i}\ \text{tail}]\ \text{list}\ \text{yellow} \unify \var{item}\ (\text{list}\ \text{blue}\ \unify \var{i}\ (\text{list}) \unify \var{tail}))$

\subitem ---
\subitem $(\text{list}\ \text{yellow}\ (\text{list}\ \text{blue}\ (\text{list})))
[\text{bound}: \text{item}\ \text{i}\ \text{tail}, \text{constrains}: \var{item} = \text{yellow}\ \var{i}=\text{blue}\ \var{tail}=(\text{list})]$
\subitem the next step is to unify all sub-tuples that are not yet checked with definitions, this are: $(\text{list}\ \text{blue}\ (\text{list}))$ and $(\text{list}))$,
\subitem $(\text{list}\ \text{blue}\ (\text{list})) \unify (list)$, fail.
\subitem $(\text{list}\ \text{blue}\ (\text{list})) \unify (\text{list}\ \var{item}\ (\text{list}))[\text{bound}: item]$
\subitem $(\text{list}\unify \text{list}\ \text{blue}\unify \var{item}\ (\text{list}) \unify (\text{list}))[\text{bound}: item]$
\subitem $(\text{list} \text{blue} (\text{list})))[\text{bound}: item, \text{constrains}: \var{item}=\text{blue}]$, succeed.
\subitem $(\text{list}\ \text{blue}\ (\text{list})) \unify (\text{list}\ \var{item}\ (\text{list}\ \var{i}\ \var{tail}))$, fail.
\subitem Finally $(list)$ will only unify with $(list)$ definition, ending the process and resulting on $\alpha = \{ (\text{list}\ \text{yellow}\ (\text{list}\ \text{blue}\ (\text{list}))) \}$
\end{enumerate}



\section{Contributors}
\label{sec:contributors}

\begin{itemize}
    \item \href{https://github.com/fsvieira}{Filipe Vieira, https://github.com/fsvieira}
\end{itemize}

\end{document}